### 2024-01-18 report/markdown/manuscript_figures_tables.Rmd
Figures, tables and methods for survival analysis

---
title: "Prognostic and biological relevance of collagen biosynthesis pathway in prostate cancer"
subtitle: "Methods, figures and tables for the transcriptome part"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: coll_biblio.bib

csl: frontiers-in-immunology.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dev = "png", 
                      dpi = 600, 
                      tab.cap.pre = 'Table ', 
                      tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Methods

## Magnetic resonance imaging of the prostate

_Dear co-authors, it's your part as experts:)_

ADC was compared between patient-matched measurements of benign prostate and malignancy by paired T test with Cohen's d effect size statistic.

## Transcriptome analyses

The transcriptome analysis was performed with R version 4.2.3. 
Its details are presented in __Supplementary Methods__.

## Transcriptome data sets

Six publicly available transcriptome data sets of primary prostate cancer samples, 
the TCGA prostate cancer cohort [@Liu2018; @Abeshouse2015] 
(n = `r tcga$expression %>% filter(tissue_type == 'tumor') %>% nrow`, RNA sequencing), 
GSE54460 [@Long2014] (n = `r gse54460$expression %>% filter(tissue_type == 'tumor') %>% nrow`, RNA sequencing), 
GSE70768  [@Ross-Adams2015] (n = `r gse70768$expression %>% filter(tissue_type == 'tumor') %>% nrow`, microarray), 
GSE70769  [@Ross-Adams2015] (n = `r gse70769$expression %>% filter(tissue_type == 'tumor') %>% nrow`, microarray), 
GSE220095 [@Schimmelpfennig2023] (n = `r gse220095$expression %>% filter(tissue_type == 'tumor') %>% nrow`, RNA sequencing), 
DKFZ [@Gerhauser2018] (n = `r dkfz$expression %>% filter(tissue_type == 'tumor') %>% nrow`, RNA sequencing) 
were re-analyzed (__Supplementary Table S1__). 
Normalized gene expression data and clinical information were fetched from the cBioportal repository (DKFZ, TCGA) or Gene Expression Omnibus (GEO). 
Selection criteria of the data sets were: availability of Gleason scoring, information of biochemical relapse and relapse-free survival, and availability of expression data for `r nrow(globals$genes_interest)` collagen-related genes covered by proteomic analysis as well (__Supplementary Table S2__).
Donor-matched benign and malignant tissue samples were available for the 
GSE70768 (n = `r norm_tumor$n_numbers$gse70768` tissue pairs) and TCGA cohort (n = `r norm_tumor$n_numbers$tcga` tissue pairs). 
Expression values were $log_2$-transformed prior to analyses. 
Non-immune cell content in cancer samples was estimated with the _MCP Counter_ and _xCell_ algorithms [@Aran2017; @Becht2016].
single sample gene set enrichment scores (ssGSEA) for the Reactome pathway gene signatures retrieved from the MSig database version 7.5.1 were computed with the _GSVA_ algorithm [@Hanzelmann2013].

The collagen-related genes were retrieved from the published Collagen Signature [@Kocher2021] and the Reactome collagen formation pathway (R-HSA-1474290) and restricted to genes covered by the proteomic analyses (__Supplementary Table S2__).

## Differential expression in benign and malignant prostate tissue and cancers stratified by Gleason score

Expression of the collagen-related genes of interest between patient-matched benign and malignant specimens was compared by the false discovery rate method (FDR) corrected paired two-tailed T test with Cohen's d effect size statistic. 
Differentially expressed genes were defined by pFDR < 0.05 and d $\geq$ 0.2 criteria indicative of significant, at least weak expression regulation (__Supplementary Table S3__).

Differential gene expression in cancer samples stratified by Gleason scoring (5 - 6, 7 and $\geq$ 8) was assessed by FDR-corrected one-way ANOVA with $\eta^2$ effect size statistic. 
Genes with pFDR < 0.05 and $\eta^2 \geq$ 0.02 were deemed differentially regulated (__Supplementary Table S4__).

## Prediction of biochemical relapse-free survival by gradient boosted machine modeling

For survival analyses, $log_2$-transformed expression values of the collagen-related genes (__Supplementary Table S2__) were subjected to batch-adjustment with the ComBat algorithm [@Leek2012]. 
Subsequently, the ComBat-adjusted expression data sets of the GSE54460, GSE70768, GSE70769, GSE220095 cohorts were merged (further referred to as 'pooled GEO' cohort).

Biochemical relapse-free survival in the pooled GEO training cohort was modeled with normalized (Z-scores) expression values of the collagen-related genes as explanatory variables by the Gradient Boosted Machine (GBM) algorithm [@Friedman2001; @Greenwell2022; @Natekin2013; @Friedman2002]. 
Selection of the optimal set of modeling parameters was accomplished by tuning with 10-fold cross-validation and out-of-fold model deviance as performance statistic (__Supplementary Tables S15__). 
The Transcriptomic Collagen Score was computed as the linear predictor score of the GBM model. 
Importance of the explanatory variables for the predictive performance of the GBM model were measured with the relative influence method [@Friedman2001] and expressed as difference in the sum of squared errors ($\Delta SSE$, __Figure \@ref(fig:fig-coll-score)A__).

Performance of the Transcriptomic Collagen Score at prediction of biochemical relapse-free survival in the training pooled GEO collective and validation data sets (TCGA and DKFZ) was assessed by univariable Cox regression with Harrell's concordance index [@Harrell1996] and integrated Brier score [@Graf1999] as metrics of goodness of fit and calibration (__Figure \@ref(fig:fig-coll-score)B__, __Supplementary Table 16__). 
Of note, predictive performance of the GBM algorithm was excellent and reproducible in the investigated cohorts as compared with several other machine learning algorithms (e.g. Ridge Cox regression, Elastic Net or Support Vector Machines, __Supplementary Figure S9__). 
Differences in biochemical relapse-free survival in cancer patients stratified by tertiles of the Transcriptomic Collagen Scores were assessed by false discovery rate (FDR) corrected Peto-Peto test (__Figure \@ref(fig:fig-coll-score)C__).

## Data and code availability

Publicly available data sets were analyzed. 
Formatted data sets used for analyses will be made available upon request to the corresponding author. 
The transcriptome R analysis pipeline is available from GitHub (https://github.com/PiotrTymoszuk/collagen_pca).

\newpage

# Figures

```{r fig-norm-tumor, fig.width = figur::convert(paper_fig$norm_tumor, to = 'in')$w, fig.height = figur::convert(paper_fig$norm_tumor, to = 'in')$h, fig.cap = 'Differences in tissue diffusion capacity and expression of collagen-related genes between the prostate cancer and benign tissue. Expression of collagen-related genes in cancers stratified by Gleason scores.'}

paper_fig$norm_tumor$plot

```

__Figure \@ref(fig:fig-norm-tumor). Differences in tissue diffusion capacity and expression of collagen-related genes between the prostate cancer and benign tissue. Expression of collagen-related genes in cancers stratified by Gleason scores.__

_(A) Representative MRI axial sequences of the prostate of an 83-year-old patient with prostate cancer (left panel). The T2-weighted images (a) show a 23mm suspicious hypo-intense lesion in the left peripheral zone (red star) of the prostate, which has an early enhancement in the contrast-enhanced T1-weighted image (b). The DWI scan (c) and ADC map (d) show a strong diffusion restriction (black and white character) of the tumor area suggesting a PI-RADS 5 lesion. Statistical significance for differences in diffusion capacity in patient-matched tumor and benign tissue measured by ADC was determined by paired T test with Cohen's d effect size statistic (right panel). Single ADC values are visualized as points, grey lines connect measurements of the same donors. The number of analyzed benign - tumor pairs, effect size and p value are displayed in the plot caption._

_(B) Differences in $log_2$-transformed expression of `r nrow(globals$genes_interest)` collagen-related genes between donor-matched tumor and benign prostate tissue samples were investigated by paired T test with Cohen's d effect size statistic in the GSE70768 (n = `r norm_tumor$n_numbers$gse70768`) and TCGA (n = `r norm_tumor$n_numbers$tcga` tissue pairs) cohorts. P values were corrected for multiple testing with the false discovery rate method. Mean normalized $log_2$-transformed expression levels in benign and cancer tissue for `r length(unique(unlist(norm_tumor$common)))` genes significantly regulated in both cohorts with at least weak effect size (pFDR < 0.05, d $\geq$ 0.2) are presented in a heat map._

_(C) Differences in $log_2$-transformed expression levels of `r nrow(globals$genes_interest)` collagen-related genes between cancer samples stratified by Gleason scores were assessed by one-way ANOVA with $\eta^2$ effect size statistic in the GSE54460 (`r format_strata_n(gs_uni$n_numbers$gse54460)`), GSE70768 (`r format_strata_n(gs_uni$n_numbers$gse70768)`), GSE70769 (`r format_strata_n(gs_uni$n_numbers$gse70769)`), GSE220095 (`r format_strata_n(gs_uni$n_numbers$gse220095)`), DKFZ (`r format_strata_n(gs_uni$n_numbers$dkfz)`), and TCGA cohort (`r format_strata_n(gs_uni$n_numbers$tcga)`). Mean normalized $log_2$-transformed expression levels of genes significantly regulated between the Gleason score strata with at least weak effect size (pFDR < 0.05, $\eta^2 \geq$ 0.02) are presented in a heat map._

\newpage

```{r fig-tme, fig.width = figur::convert(paper_fig$tme, to = 'in')$w, fig.height = figur::convert(paper_fig$tme, to = 'in')$h, fig.cap = 'Expression of collagen-related genes in components of the tumor microenvironment.'}

paper_fig$tme$plot

```

__Figure \@ref(fig:fig-tme). Expression of collagen-related genes in components of the tumor microenvironment.__

_(A) Cancer samples in six published cohorts were classified as collagen low and collagen high tumors in respect to `r nrow(globals$genes_interest)` collagen-related transcripts by semi-supervised PAM clustering (GSE54460: `r format_strata_n(ana_globals$n_numbers$gse54460)`, GSE70768: `r format_strata_n(ana_globals$n_numbers$gse70768)`, GSE70769: `r format_strata_n(ana_globals$n_numbers$gse70769)`, GSE220095: `r format_strata_n(ana_globals$n_numbers$gse220095)`, DKFZ: `r format_strata_n(ana_globals$n_numbers$dkfz)`, TCGA: `r format_strata_n(ana_globals$n_numbers$tcga)`). Counts of cancer-associated fibroblasts and endothelial cells were predicted by the MCP Counter algorithm and compared between the collagen clusters by Mann-Whitney test with r effect size statistic. Median normalized cell counts with interquartile ranges are presented as boxes, whiskers span over 150% of the interquartile ranges. Values for single cancer samples are depicted as points. All presented effects were significant (pFDR < 0.05) with at least weak (endothelial cells, r > 0.21) or moderate effect size (fibroblasts, r > 0.55)._

_(B) Placeholder for the single cell RNA seq data._

\newpage

```{r fig-coll-score, fig.width = figur::convert(paper_fig$coll_score, to = 'in')$w, fig.height = figur::convert(paper_fig$coll_score, to = 'in')$h, fig.cap = 'Transcriptional Collagen Model of biochemical relapse-free survival.'}

paper_fig$coll_score$plot

```

__Figure \@ref(fig:fig-coll-score). Transcriptional Collagen Model of biochemical relapse-free survival.__

_Transcriptional Collagen Score was established in the pooled GEO training cohort by multi-parameter Gradient Boosted Machine (GBM) modeling of biochemical relapse-free survival as a function of normalized $log_2$-transformed and ComBat-adjusted expression levels of `r nrow(globals$genes_interest)` collagen-related genes._

_(A) Importance of explanatory variables measured by the gradient of sum of squared errors ($\Delta SSE$) during the training process attributed to particular collagen-related transcript levels._

_(B) Performance of the Elastic Net Cox model at prediction of biochemical relapse-free survival in the training TCGA cohort and test collectives assessed by concordance index (C-index, high values indicate high concordance between the predicted and observed survival) and integrated Brier score (IBS, low values indicate good model fit and proper calibration). Values of C-index and IBS expected for random survival prediction are depicted as dashed lines._

_(C) Fractions of biochemical relapse-free patients in low, intermediate and high tertiles of Transcriptional Collagen Score in the training TCGA cohort, and the GSE70768 and DKFZ validation collectives presented in Kaplan-Meier plots. Differences in survival between the tertiles were investigated by Peto-Peto test. Total numbers of patients and numbers of biochemical relapses are displayed in the plot captions. P values are shown in the plots._

\newpage

# References